---
title: "Class 10: Halloween Candy"
format: gfm
date: 10/28/22
---

In this mini-project we will examine 538 Halloween Candy data. 

```{r}
candy_file <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv"

```

```{r}
candy = read.csv(candy_file,row.names=1)
head(candy)
```

Q1. How many differeny candy types are in this dataset?
```{r}
nrow(candy)
```

Q2. How many fruity candy types are in this dataset?
```{r}
sum(candy$fruity)
```
Q3. What is your favorite candy in the dataset and what is it's `winpercent` value?

```{r}
candy["Kit Kat", ]$winpercent
```

The winpercent for Kit Kat is `r candy["Kit Kat", ]$winpercent`

Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat", ]$winpercent
```

The winpercent for Kit Kat is `r candy["Kit Kat", ]$winpercent`

Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```

```{r}
skimr::skim(candy)
```

Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

The winpercent variable seems to be on a different scale than the rest. The rest of the values are functions

Q8. Plot a histogram of winpercent values

```{r}
library(ggplot2)
ggplot(candy)+ 
  aes(winpercent) +
  geom_histogram(bins=10,col="black", fill="orange")
```

Q9. Is the distribution of winpercent values symmetrical?
No it's not

Q10. Is the center of the distribution above or below 50%?
It is below 50%

Q11. On average is chocolate candy higher or lower ranked than fruit candy?

Let's look at chocolate
```{r}
chocolate_inds <- as.logical(candy$chocolate)
chocolate_win <- candy$winpercent[chocolate_inds]
mean(chocolate_win)

```

Now let's look at fruit candy..
```{r}
fruity_inds <- as.logical(candy$fruity)
fruity_win <- candy$winpercent[fruity_inds]
mean(fruity_win)
```
On average, chocolate is ranked higher than fruit candy.

Q12. Is this difference statistically significant?

```{r}
t.test(chocolate_win, fruity_win)
```
According to the results of the t-test, the difference is statistically different 

Q13. What are the five least liked candy types in this set?

```{r}
library(dplyr)

candy%>%
  arrange(winpercent)%>%
  head(5)
```

The five least liked candy types are Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble, and Jawbusters

Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
library(dplyr)

candy%>%
  arrange(winpercent)%>%
  tail(5)
```
The top 5 favorite candies are Snickers, Kit Kat, Twix, ReeseÕs Miniatures, and ReeseÕs Peanut Butter cup

Q15. Make a first barplot of candy ranking based on winpercent values.
```{r}
library(ggplot2)

ggplot(candy) +
  aes(winpercent, rownames(candy)) + 
  geom_col()
```

Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
library(ggplot2)

ggplot(candy) +
  aes(winpercent, reorder(rownames(candy),winpercent)) + 
  geom_col()
```
```{r}
ggsave("myrplot.png", height=15)
```

Adding color vectors
```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "blue"
my_cols[as.logical(candy$fruity)] = "red"
```

Add it to the graph

```{r}
ggplot(candy)+ aes(winpercent, reorder(rownames(candy), winpercent))+ geom_col(fill=my_cols)
```

Q17. What is the worst ranked chocolate candy?

Sixlets it the worst ranked chocolate candy

Q18. What is the best ranked fruity candy?

Starburst

# Looking at pricepercent

What about value for money?
What is the best candy for the least money?

One way to get at this would be to make a plot of `winpercent` vs `pricepercent`
```{r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 7)
```
Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

The candy that is the highest ranked for the least money is Reeses Miniatures


Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
library(dplyr)
candy%>%
  arrange(pricepercent)%>%
  tail(5)
```

```{r}
# Make a lollipop chart of pricepercent
ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), 
                   xend = 0), col="gray40") +
    geom_point()
```
Exploring Correlation

```{r}
library(corrplot)
```

```{r}
cij <-  cor(candy)
corrplot(cij)
```

Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

Fruity and chocolate are anti-correlated

Q23. Similarly, what two variables are most positively correlated?

bar and chocolate are most positively correlated

# Principal Component Analysis

```{r}
pca <- prcomp(candy, scale=T)
summary(pca)
```

Let's plot results from PCA analysis

```{r}
plot(pca$x[,1:2])
```
Adding color we originally defined

```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```

```{r}
my_data <- cbind(candy, pca$x[,1:3])
```

```{r}
p <- ggplot(my_data)+
  aes(x=PC1, y=PC2, 
      size=winpercent/100,
      text=rownames(my_data),
      label=rownames(my_data))+
  geom_point(col=my_cols)
p
```

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (blue), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

```{r}
library(plotly)
```

```{r}
# ggplotly(p)
```

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

The original variable that are picked up most strongly by PC1 in the positive direction are fruity, hard, and pluribus

These do make sense because the candies that are furthest to the right on PC1 are fruity, hard, and a lot of them do come in multiples in a packet
